---
title: "Interactive Visualization with R"
author: '[Christopher B. Cole](https://github.com/Chris1221)'
date: "Monday August 24th 2015 <br> Peers Teaching Peers Tutorial Series"
output:
  ioslides_presentation:
    kepp_html: yes
    keep_md: yes
    logo: logo.jpeg
runtime: shiny
subtitle: ' New Opportunities and Old Challenges'
resource_files:
- logo.jpeg
---

## Outline

1. Introduction to Interactive Visualization
2. Requirements
3. Steps to Successful Visualization
4. Examples
5. Live Demo


# 1. Intro to Interactive Visualization


## Interactive Visualization

- Explore high dimensional data
    - See trends in real time
- Better understand and manipulate data
    - Engage audiences and encourage participation
- Report, share, and publicize 
    - Reach a wider audience with easier interpretation of figures

## Why `R`?

- Turing complete language
    - Anything you could ever want to do
- Works with existing analysis pipelines
- Easy conversion from regular plots 
    - `ggplot2`, etc.
- Easy dissemination and publication

# 2. Requirements

## Software

- `R` + `RStudio`
    - IDE with integrated `html` viewer, output control, and so much more. 
- `Shiny` + `RPubs` + `GitHub`
    - Source control + Publication
- Markdown + RMarkdown + Pandocs + `Knitr`
    - Easy way to create a website with all the correct rendering 
- Packages
    - `ggvis`, `ggplot2`, `rCharts`, `htmlwidgets`, `plot.ly`, `googleVis`, etc...

## Concepts

- Three basic things
    - Web server 
    - Code 
    - Data
- Github repository

# 3. Steps to Successful Visualization

## Thought process

1. What is the message that you want to show?
2. How do you want to show it?
3. How can available tools help you?
4. How do I use these tools?
5. How do I share my research?

## Practical Thinking

1. Get your data in line
    - `tidyr`, `Excel`, etc.
2. Figure out how you want to display your data
    - What have similar publications used? 
    - How can you improve on this?
3. Code it
    - **BE REPRODUCIBLE**
    - Save your code on `Github` etc.
4. Share it
    - Learn how to show people your data
    - `Shiny` + `Rpubs` etc.

# Examples

## ggvis

- Updated `ggplot2` with interactive features
- Developed by Hadley Wickham
- Easy, simple syntax
- Almost universally usable 

## ggvis Example

```{r, echo=FALSE}
library(ggvis)

mtcars %>%
  ggvis(~wt, ~mpg) %>%
  layer_smooths(span = input_slider(0.5, 1, value = 1)) %>%
  layer_points(size := input_slider(100, 1000, value = 100))
```

## ggvis Example

```{r, echo=TRUE, eval = F}
library(ggvis)

mtcars %>%
  ggvis(~wt, ~mpg) %>%
  layer_smooths(span = input_slider(0.5, 1, value = 1)) %>%
  layer_points(size := input_slider(100, 1000, value = 100))
```

## R Tips and Tricks

- `dplyr::%>%`
    - Pipe operator, similar to *nux `|`
      - `cat file.txt | wc -l`
- Example
    - `"apple" %>% wash() %>% cut() %>% eat()`

~~~~~~~
    food <- "apple"
    washed <- wash(food)
    cut_up <- cut(washed)
    eaten <- eat(apple)
~~~~~~~

## plot.ly

- Startup from Montreal
- Simple online interface 
- API for more complex use
- Fast, easy, customizable. 
- Premium = Private
- Publication quality images even as static

## plot.ly

<iframe scrolling='no' seamless='seamless' style='border:none'
src='https://plot.ly/~MattSundquist/2509' width='800' height='500'></iframe>

## plot.ly

<iframe scrolling='no' seamless='seamless' style='border:none'
src='https://plot.ly/~ChrisCole/40/average-difference-in-true-positives-with-maf/' width='800' height='500'></iframe>

## plot.ly

`<iframe scrolling='no' seamless='seamless' style='border:none'`
`src='https://plot.ly/~MattSundquist/2509' width='800' height='500'></iframe>`

## Anything else really

<iframe scrolling='no' seamless='seamless' style='border:none'
src='https://shamiri.shinyapps.io/NRLdashboard/' width='800' height='500'></iframe>

## htmlwidgets

- Ramnath Vaidyanathan from McGill
- Port of D3.js (javascript)
- Suite of packages
- Still heavily under development
- Very easy to use

## htmlwidgets

```{r, echo = FALSE}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```

## htmlwidgets 2

```{r, echo = F}
library(networkD3)
data(MisLinks, MisNodes)
forceNetwork(Links = MisLinks, Nodes = MisNodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             Group = "group", opacity = 0.4)
```

## htmlwidgets 3

```{r, echo = F, message = F}
library(DT)
datatable(iris, options = list(pageLength = 5))
```

## googleVis

- `R` interface to Google Charts API
- Very easy to use
- Attractive
- Google Maps etc. 
- Not really compatible with Shiny

## googleVis

```{r, echo = F, message = F}
library(googleVis)
op <- options(gvis.plot.tag="chart")
```

```{r results='asis', tidy=TRUE, echo = F}

mark_func <- function(d) {
    d$Mean=mean(d$Popularity)
    CC <- gvisComboChart(d, xvar='City',
          yvar=c('Mean', 'Popularity'),
          options=list(seriesType='bars',
                       width=450, height=300,
                       title='City Popularity',
                       series='{0: {type:\"line\"}}'))
    return(CC)
}
```

```{r results='asis', tidy=TRUE}
plt <- mark_func(CityPopularity)
plot(plt)
```  

# 5. Live demo...

# Conclusion

## Materials

- Slideshow
    - Code: `https://github.com/Chris1221/PTP_presentation_2015`
    - Rendered Slideshow: https://ccole.shinyapps.io/PTP_presentation
- Shiny Cheat Sheet: `http://shiny.rstudio.com/articles/cheatsheet.html`
- Markdown Cheat Sheet: `https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet`
- More Materials: `https://github.com/Chris1221/PTP_presentation_2015`
